---
title: "01_exploration"
output: html_document
date: "2025-04-01"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Load libraries
```{r}

library(tidyverse)

```
Read lahman database
```{r}
Teams <- read_csv("data/lahman_1871-2023_csv/Teams.csv")
People <- read_csv("data/lahman_1871-2023_csv/People.csv")
TeamsFranchises <- read_csv("data/lahman_1871-2023_csv/TeamsFranchises.csv")
Parks <- read_csv("data/lahman_1871-2023_csv/Parks.csv")
Batting <- read_csv("data/lahman_1871-2023_csv/Batting.csv")
Pitching <- read_csv("data/lahman_1871-2023_csv/Pitching.csv")
Fielding <- read_csv("data/lahman_1871-2023_csv/Fielding.csv")
FieldingOF <- read_csv("data/lahman_1871-2023_csv/FieldingOF.csv")
FieldingOFsplit <- read_csv("data/lahman_1871-2023_csv/FieldingOFsplit.csv")
Appearances <- read_csv("data/lahman_1871-2023_csv/Appearances.csv")
Managers <- read_csv("data/lahman_1871-2023_csv/Managers.csv")
#... <- read_csv("data/lahman_1871-2023_csv/...")
#... <- read_csv("data/lahman_1871-2023_csv/...")



```

Explore People table

How many players with the given name "Bob" have ever played?
List of Bobs, arrange by debut date. Showing ID, name, birth and death years, debut and finalgame.
```{r}
People |>
  filter(nameFirst == "Bob") |>
  arrange(debut, finalGame) |>
  select(playerID, nameFirst, nameLast, birthYear, debut, finalGame, deathYear)
  
```

Has there ever been a Bob Roberts?
```{r}

People |>
  filter(nameFirst == "Bob" & str_detect(nameLast, "Robert")) |>
    select(playerID, nameFirst, nameLast, birthYear, debut, finalGame, deathYear)


```

How about a Bob Bob-something?
```{r}

People |>
  filter(nameFirst == "Bob" & str_detect(nameLast, "Bob"))
```

Group Bobs by decade of debut

```{r}
People |>
  select(playerID, nameFirst, nameLast, birthYear, debut, finalGame, deathYear) |>
  filter(nameFirst == "Bob") |>
  group_by(debut) |>
  arrange(debut)

```

```{r}
People |>
  select(playerID, nameFirst, nameLast, birthYear, debut, finalGame, deathYear) |>
  filter(nameFirst == "Bob") |>
  group_by(debut) |>
  arrange(debut) |>
  filter(str_detect(debut, "1872"))
```

Split debut year into new column
```{r}
People |>
  select(playerID, nameFirst, nameLast, birthYear, debut, finalGame, deathYear) |>
  filter(nameFirst == "Bob") |>
  arrange(debut) |>
  separate(debut, sep="-", into=c("debut_yr", "debut_mo", "debut_day"))
  



```

Group by debut year
```{r}
Bob_debut_by_yr <- People |>
  select(playerID, nameFirst, nameLast, birthYear, debut, finalGame, deathYear) |>
  filter(nameFirst == "Bob") |>
  arrange(debut) |>
  separate(debut, sep="-", into=c("debut_yr", "debut_mo", "debut_day")) |>
  group_by(debut_yr) |>
  summarize(n=n())
```

Visualize Bob debuts by year
```{r}
People |>
  select(playerID, nameFirst, nameLast, birthYear, debut, finalGame, deathYear) |>
  filter(nameFirst == "Bob") |>
  arrange(debut) |>
  separate(debut, sep="-", into=c("debut_yr", "debut_mo", "debut_day")) |>
  ggplot(aes(x = debut_yr)) +
  geom_bar()

```

Arrange years by most Bob debuts
```{r}
Bob_debut_by_yr |>
  arrange(desc(n))
```

